---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.18.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Feature Selection 
## Multiple Methods: LASSO, Elastic Net, Random Forest, Correlation Analysis


## 1. Load Libraries

```{r vscode={'languageId': 'r'}}
library(glmnet)
library(survey)
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(readr)
library(randomForest)
library(caret)
library(gridExtra)
```

## Load Data

```{r vscode={'languageId': 'r'}}
# Load cleaned NHANES data and survey design object
load("./data/cleaned_nhanes_data.RData")

# Load variable labels
var_labels <- read_csv("./data/variable_labels.csv")


```

## Prepare Features and Outcome
### Define outcome variable and predictors

```{r vscode={'languageId': 'r'}}
# Define outcome variable 
#outcome_var <- "LBXTST"  # Testosterone
#outcome_var <- "LBXEST"  # Estradiol 
outcome_var <- "LBXSHBG" # SHBG 

# Define weight variable
weight_var <- "WTSB2YR"

# Variables to exclude from predictors
exclude_vars <- c(
  "SEQN",           # ID
  "WTMEC2YR",       # Weights
  "WTSSBH2Y",
  "WTSB2YR",
  "SDMVPSU",        # Survey design variables
  "SDMVSTRA",
  "RIDEXPRG",       # Constant after filtering (all non-pregnant)
  "RIAGENDR",       # Constant after filtering (all female)
  "LBXTST",         # Outcome variables (exclude all)
  "LBXEST",
  "LBXSHBG",
  "LBDTSTLC",       # Comment codes
  "LBDESTLC",
  "LBDSHGLC"
)

# Get all EDC (chemical) variable names
edc_vars <- names(data_1)[grepl("^URX|^SSU", names(data_1))]
edc_vars <- setdiff(edc_vars, c("URXUCR"))  # Keep creatinine as covariate

# Remove comment codes (ending in LC, PLC, etc.)
edc_vars <- edc_vars[!grepl("LC$|PLC$", edc_vars)]

cat("Number of EDC variables:", length(edc_vars), "\n")
cat("EDC variables:\n")
print(edc_vars)
```

### Log-transform chemical variables and prepare data matrix

```{r vscode={'languageId': 'r'}}
# Create log-transformed versions of EDC variables
data_transformed <- data_1

for (var in edc_vars) {
  new_var <- paste0("log_", var)
  data_transformed[[new_var]] <- log(data_transformed[[var]] + 1)
}

# Get log-transformed EDC variable names
log_edc_vars <- paste0("log_", edc_vars)

# Select predictor variables: log-transformed EDCs + other covariates
other_covariates <- c("RIDAGEYR", "DMDEDUC2", "INDFMPIR", "URXUCR", "BMXBMI", 
                     "RHQ031", "RHQ074", "RHQ076", "RHQ131", "RHQ160", 
                     "RHQ171", "RHD280", "RHQ078", "RHD180",
                     "ALQ120Q", "SMQ040", "SMD650")

# Filter to only variables that exist in the data
other_covariates <- other_covariates[other_covariates %in% names(data_transformed)]

predictor_vars <- c(log_edc_vars, other_covariates)

cat("Total number of predictors:", length(predictor_vars), "\n")
cat("  - Log-transformed EDCs:", length(log_edc_vars), "\n")
cat("  - Other covariates:", length(other_covariates), "\n")
```

### Create predictor matrix and outcome vector

```{r vscode={'languageId': 'r'}}
# Remove rows with missing outcome
complete_data <- data_transformed[!is.na(data_transformed[[outcome_var]]), ]

# Extract outcome
outcome <- complete_data[[outcome_var]]

# Extract weights
weights <- complete_data[[weight_var]]

# Create predictor matrix (handle missing values)
predictors_df <- complete_data[, predictor_vars]

# Impute missing values with median for each column
for (col in names(predictors_df)) {
  if (any(is.na(predictors_df[[col]]))) {
    predictors_df[[col]][is.na(predictors_df[[col]])] <- median(predictors_df[[col]], na.rm = TRUE)
  }
}

# Convert to matrix
predictors <- as.matrix(predictors_df)

cat("Final dimensions:\n")
cat("  Predictors:", dim(predictors), "\n")
cat("  Outcome:", length(outcome), "\n")
cat("  Weights:", length(weights), "\n")
cat("\nOutcome summary (", outcome_var, "):\n", sep="")
print(summary(outcome))
```

## Feature Selection

### Method 1: LASSO Regression

```{r vscode={'languageId': 'r'}}
cat("\n", paste(rep("=", 60), collapse=""), "\n")
cat("LASSO Feature Selection\n")
cat(paste(rep("=", 60), collapse=""), "\n\n")

# LASSO with cross-validation
set.seed(42)
cv_lasso <- cv.glmnet(
  x = predictors,
  y = outcome,
  weights = weights,
  alpha = 1,  # 1 = LASSO
  standardize = TRUE,
  nfolds = 10
)

# Plot cross-validation curve
plot(cv_lasso, main = "LASSO Cross-Validation")

# Get coefficients at lambda.min and lambda.1se
coef_min <- coef(cv_lasso, s = "lambda.min")
coef_1se <- coef(cv_lasso, s = "lambda.1se")

# Extract non-zero coefficients (lambda.min)
lasso_selected_min <- data.frame(
  feature = rownames(coef_min)[coef_min[,1] != 0],
  coefficient = coef_min[coef_min[,1] != 0, 1]
)
lasso_selected_min <- lasso_selected_min[lasso_selected_min$feature != "(Intercept)", ]
lasso_selected_min$abs_coef <- abs(lasso_selected_min$coefficient)
lasso_selected_min <- lasso_selected_min[order(-lasso_selected_min$abs_coef), ]

cat("\nLambda min:", cv_lasso$lambda.min, "\n")
cat("Lambda 1se:", cv_lasso$lambda.1se, "\n")
cat("\nNumber of selected features (lambda.min):", nrow(lasso_selected_min), "\n\n")
cat("Top 20 LASSO selected features:\n")
print(head(lasso_selected_min, 20))
```

### Method 2: Elastic Net

```{r vscode={'languageId': 'r'}}
cat("\n", paste(rep("=", 60), collapse=""), "\n")
cat("Elastic Net Feature Selection (alpha = 0.5)\n")
cat(paste(rep("=", 60), collapse=""), "\n\n")

# Elastic Net with cross-validation (alpha = 0.5, midway between LASSO and Ridge)
set.seed(42)
cv_enet <- cv.glmnet(
  x = predictors,
  y = outcome,
  weights = weights,
  alpha = 0.5,  # 0.5 = Elastic Net
  standardize = TRUE,
  nfolds = 10
)

# Plot cross-validation curve
plot(cv_enet, main = "Elastic Net Cross-Validation")

# Get coefficients
coef_enet <- coef(cv_enet, s = "lambda.min")

# Extract non-zero coefficients
enet_selected <- data.frame(
  feature = rownames(coef_enet)[coef_enet[,1] != 0],
  coefficient = coef_enet[coef_enet[,1] != 0, 1]
)
enet_selected <- enet_selected[enet_selected$feature != "(Intercept)", ]
enet_selected$abs_coef <- abs(enet_selected$coefficient)
enet_selected <- enet_selected[order(-enet_selected$abs_coef), ]

cat("\nNumber of selected features:", nrow(enet_selected), "\n\n")
cat("Top 20 Elastic Net selected features:\n")
print(head(enet_selected, 20))
```

### Method 3: Random Forest Feature Importance

```{r vscode={'languageId': 'r'}}
cat("\n", paste(rep("=", 60), collapse=""), "\n")
cat("Random Forest Feature Importance\n")
cat(paste(rep("=", 60), collapse=""), "\n\n")

# Random Forest with weighted samples
set.seed(42)
rf_model <- randomForest(
  x = predictors,
  y = outcome,
  ntree = 500,
  importance = TRUE,
  mtry = floor(sqrt(ncol(predictors))),
  nodesize = 5
)

# Get variable importance
importance_scores <- importance(rf_model)
rf_importance <- data.frame(
  feature = rownames(importance_scores),
  IncMSE = importance_scores[, "%IncMSE"],
  IncNodePurity = importance_scores[, "IncNodePurity"]
)
rf_importance <- rf_importance[order(-rf_importance$IncMSE), ]

cat("\nRandom Forest Performance:\n")
cat("  R-squared:", round(tail(rf_model$rsq, 1), 4), "\n")
cat("  MSE:", round(tail(rf_model$mse, 1), 2), "\n\n")
cat("Top 20 important features by %IncMSE:\n")
print(head(rf_importance, 20))

# Variable importance plot
varImpPlot(rf_model, n.var = 20, main = "Top 20 Random Forest Variable Importance")
```

### Method 4: Correlation Analysis

```{r vscode={'languageId': 'r'}}
cat("\n", paste(rep("=", 60), collapse=""), "\n")
cat("Correlation-based Feature Selection\n")
cat(paste(rep("=", 60), collapse=""), "\n\n")

# Calculate correlation with outcome
correlations <- apply(predictors, 2, function(x) cor(x, outcome, use = "complete.obs"))

corr_df <- data.frame(
  feature = names(correlations),
  correlation = correlations,
  abs_correlation = abs(correlations)
)
corr_df <- corr_df[order(-corr_df$abs_correlation), ]

cat("Top 20 features by absolute correlation with outcome:\n")
print(head(corr_df, 20))
```

## 5. Compare Methods and Find Consensus Features

```{r vscode={'languageId': 'r'}}
cat("\n", paste(rep("=", 60), collapse=""), "\n")
cat("Comparing Feature Selection Methods\n")
cat(paste(rep("=", 60), collapse=""), "\n\n")

# Get ALL selected features from each method (not just top N)
lasso_all <- lasso_selected_min$feature
enet_all <- enet_selected$feature
# For Random Forest, take features with positive importance (or use a threshold)
rf_threshold <- 0  # Include all features with positive importance
rf_all <- rf_importance$feature[rf_importance$IncMSE > rf_threshold]
# For correlation, use a significance threshold (e.g., |r| > 0.05 or top features)
corr_threshold <- 0.05
corr_all <- corr_df$feature[corr_df$abs_correlation > corr_threshold]

cat("Number of features selected by each method:\n")
cat("  LASSO:", length(lasso_all), "\n")
cat("  Elastic Net:", length(enet_all), "\n")
cat("  Random Forest (IncMSE > ", rf_threshold, "):", length(rf_all), "\n", sep="")
cat("  Correlation (|r| > ", corr_threshold, "):", length(corr_all), "\n\n", sep="")

# Combine all selected features
all_selected <- unique(c(lasso_all, enet_all, rf_all, corr_all))

# Count how many methods selected each feature
consensus <- data.frame(
  feature = all_selected,
  lasso = all_selected %in% lasso_all,
  elastic_net = all_selected %in% enet_all,
  random_forest = all_selected %in% rf_all,
  correlation = all_selected %in% corr_all
)

consensus$method_count <- rowSums(consensus[, c("lasso", "elastic_net", "random_forest", "correlation")])
consensus <- consensus[order(-consensus$method_count), ]

cat("Total unique features across all methods:", nrow(consensus), "\n\n")
cat("Consensus features (selected by multiple methods):\n")
print(consensus)

# Features selected by all methods
all_methods <- consensus[consensus$method_count == 4, ]
cat("\n✓ Features selected by ALL 4 methods (", nrow(all_methods), "):\n", sep="")
if (nrow(all_methods) > 0) {
  for (feat in all_methods$feature) {
    cat("  -", feat, "\n")
  }
} else {
  cat("  None\n")
}

# Features selected by at least 3 methods
three_methods <- consensus[consensus$method_count >= 3, ]
cat("\n✓ Features selected by at least 3 methods (", nrow(three_methods), "):\n", sep="")
if (nrow(three_methods) > 0) {
  for (feat in three_methods$feature) {
    cat("  -", feat, "\n")
  }
} else {
  cat("  None\n")
}

# Features selected by at least 2 methods
two_methods <- consensus[consensus$method_count >= 2, ]
cat("\n✓ Features selected by at least 2 methods (", nrow(two_methods), "):\n", sep="")
if (nrow(two_methods) > 0) {
  for (feat in two_methods$feature) {
    cat("  -", feat, "\n")
  }
}
```

## 6. Visualizations


### Plot 1: LASSO Coefficient Plot

```{r vscode={'languageId': 'r'}}
# LASSO coefficients
top_lasso <- head(lasso_selected_min, 20)

ggplot(top_lasso, aes(x = reorder(feature, coefficient), y = coefficient)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 20 Features - LASSO Coefficients",
       x = "Feature",
       y = "Coefficient") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))
```

### Random Forest Importance

```{r vscode={'languageId': 'r'}}
# Random Forest importance
top_rf <- head(rf_importance, 20)

ggplot(top_rf, aes(x = reorder(feature, IncMSE), y = IncMSE)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Top 20 Features - Random Forest Importance",
       x = "Feature",
       y = "% Increase in MSE") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))
```

### Method Comparison Heatmap

```{r vscode={'languageId': 'r'}}

```

```{r vscode={'languageId': 'r'}}

```

```{r vscode={'languageId': 'r'}}
# Create binary matrix for heatmap
top_consensus <- head(consensus, 30)
heatmap_data <- as.matrix(top_consensus[, c("lasso", "elastic_net", "random_forest", "correlation")])
rownames(heatmap_data) <- top_consensus$feature

# Convert logical to numeric
heatmap_data <- heatmap_data * 1

# Create heatmap
library(pheatmap)
pheatmap(
  heatmap_data,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = c("white", "#d73027"),
  legend_breaks = c(0, 1),
  legend_labels = c("Not Selected", "Selected"),
  main = "Feature Selection Across Methods",
  fontsize_row = 7,
  fontsize_col = 9
)
```



<!-- #region vscode={"languageId": "r"} -->

<!-- #endregion -->

```{r vscode={'languageId': 'r'}}

```

```{r vscode={'languageId': 'r'}}

```



<!-- #region vscode={"languageId": "r"} -->
# Summary of Feature Selection Results for Estradiol

## Method Performance

### LASSO Regression
- **Selected features:** 0 features at lambda.min
- **Issue:** The regularization was too strong, shrinking all coefficients to zero
- **Interpretation:** This suggests weak linear relationships or high multicollinearity

### Elastic Net (alpha = 0.5)
- **Selected features:** Very few or no features (likely similar to LASSO)
- **Issue:** Similar problem to LASSO
- **Interpretation:** Linear methods struggled to identify meaningful predictors

### Random Forest
- **Selected features:** 15 important features based on variable importance
- **Top features included:** Various phthalate metabolites and covariates
- **Interpretation:** Shows non-linear relationships that LASSO/Elastic Net missed

### Correlation Analysis
Top correlations with estradiol (though all relatively weak):

| Variable | Description | Correlation (r) | Direction |
|----------|-------------|-----------------|-----------|
| RIDAGEYR | Age | 0.158 | Positive |
| RHQ131 | Reproductive history | -0.113 | Negative |
| URXUCR | Creatinine | 0.098 | Positive |
| log_SSURHIBP | BPA metabolite | 0.078 | Positive |
| Various phthalates | Phthalate metabolites | 0.05-0.07 | Positive |

---

## Key Findings

### Method Consensus
- **Selected by all 4 methods:** None
- **Selected by 3+ methods:** None
- **Selected by 2 methods** (Random Forest + Correlation):
  - `log_URXMBP`
  - `URXUCR`
  - `log_SSURHIBP`
  - `RIDAGEYR`
  - `log_URXMIB`
  - `log_URXMHH`
  - `log_URXPPB`

---

## Important Covariates

### Demographic Factors
- **Age (RIDAGEYR):** Strongest predictor
  - Relationship: Older age → Higher estradiol
  - Correlation: r = 0.158

### Reproductive History
- **RHQ131** (Ever been pregnant)
- **RHD280** (Had hysterectomy)
- **RHD180** (Age at first live birth)

### Anthropometric Measures
- **BMI:** Negative correlation
  - Relationship: Higher BMI → Lower estradiol

---

## EDC Chemicals of Interest

### Bisphenols
- **SSURHIBP** (BPA metabolite)

### Phthalate Metabolites
- `URXMHP`
- `URXMIB`
- `URXMHH`
- `URXPPB`
- `URXECP`
- `URXMBP`

---

## Overall Interpretation

> **The relationships between EDCs and estradiol appear to be weak and non-linear**, which explains why linear methods (LASSO/Elastic Net) failed to identify features while Random Forest found some signal.

### Implications for Modeling
1. **Linear models may not be appropriate** for capturing EDC-hormone relationships
2. **Non-linear methods** (Random Forest, GAM) may be more suitable
3. **Weak effect sizes** suggest need for larger sample sizes or more sensitive measures
4. **High multicollinearity** among EDCs requires careful model interpretation
5. **Mixture models** (WQS, BKMR) may be needed to capture combined effects

### Next Steps
- Consider non-linear modeling approaches (GAM, Random Forest regression)
- Implement mixture analysis methods
- Stratify by age groups or reproductive status
- Adjust for multiple testing in final models
<!-- #endregion -->

## Summary Report - LBXTST


# Summary of Feature Selection Results for Testosterone

## Method Performance

### LASSO Regression
- **Selected features:** 1 feature
  - `URXMEP` (Mono-ethyl phthalate)
- **Note:** Questionable credibility due to selecting only a single feature
- **Interpretation:** Strong regularization selected only the most dominant predictor

### Elastic Net (alpha = 0.5)
- **Selected features:** 1 feature (same as LASSO)
  - `URXMEP` (Mono-ethyl phthalate)
- **Interpretation:** Consistent with LASSO, indicating this is the strongest linear predictor

### Random Forest
- **Selected features:** 15 important features based on variable importance
- **Notable:** Did NOT identify URXMEP as important
- **Interpretation:** Identified different features based on non-linear relationships and interactions

### Correlation Analysis
- **Selected features:** 15 features with strongest linear relationships
- **Top categories:**
  - Phthalates: URXMEP, URXMPB, URXEPB
  - Smoking behavior
  - Reproductive history variables

---

## Key Findings

### Method Consensus
- **Selected by 2+ methods:**
  - `log_URXCOP` (Mono(carboxyoctyl) phthalate): Random Forest + Correlation
  - `log_URXMPB` (Methyl paraben): Random Forest + Correlation
  - `log_URXEPB` (Ethyl paraben): Random Forest + Correlation

### Most Promising EDC
**URXMEP (Mono-ethyl phthalate)**
- **Correlation:** r = 0.19 (strongest among all chemicals)
- **Selected by:** LASSO, Elastic Net, Correlation analysis
- **NOT selected by:** Random Forest (interesting discrepancy)

---

## Top Correlations with Testosterone

| Variable | Description | Correlation (r) | Direction |
|----------|-------------|-----------------|-----------|
| **URXMEP** | **Mono-ethyl phthalate** | **0.19** | **Positive** |
| SMQ040 | Current smoking status | 0.13 | Positive |
| RHQ171 | Number of live births | -0.12 | Negative |
| RHQ160 | Number of pregnancies | -0.11 | Negative |
| RIDAGEYR | Age | -0.08 | Negative |

---

## Important Covariates

### Lifestyle Factors
- **SMQ040** (Current smoking status)
  - Correlation: r = 0.13
  - Relationship: Smoking → Higher testosterone

### Reproductive History
- **RHQ171** (Number of live deliveries)
  - Correlation: r = -0.12
  - Relationship: More births → Lower testosterone
- **RHQ160** (Number of pregnancies)
  - Correlation: r = -0.11
  - Relationship: More pregnancies → Lower testosterone
- **RHQ031** (Regular menstrual periods in last 12 months)
  - Identified by correlation analysis

### Demographic Factors
- **RIDAGEYR** (Age)
  - Correlation: r = -0.08
  - Relationship: Older age → Lower testosterone

---

## EDC Chemicals of Interest

### Phthalate Metabolites
- **URXMEP** (Mono-ethyl phthalate) - **Strongest signal (r = 0.19)**
- `log_URXCOP` (Mono(carboxyoctyl) phthalate)

### Parabens
- `log_URXMPB` (Methyl paraben)
- `log_URXEPB` (Ethyl paraben)

---

## Overall Interpretation

> **URXMEP (mono-ethyl phthalate) emerges as the most promising EDC biomarker for testosterone**, with the strongest correlation (r = 0.19) and selection by multiple linear methods.

### Key Observations

1. **Linear vs. Non-linear Methods Disagreement:**
   - LASSO/Elastic Net: Selected URXMEP only
   - Random Forest: Found 15 features but missed URXMEP
   - Suggests URXMEP has strong linear relationship, while other chemicals may have non-linear effects

2. **Stronger Relationships than Estradiol:**
   - Testosterone shows stronger correlations with EDCs (up to r = 0.19)
   - More features selected by LASSO/Elastic Net
   - Clearer linear relationships

3. **Reproductive History Matters:**
   - Number of pregnancies and live births negatively associated with testosterone
   - Hormonal changes from pregnancy may have lasting effects

4. **Smoking as Confounder:**
   - Positive association with testosterone
   - Important covariate to adjust for in models

### Implications for Modeling

1. **Primary focus:** URXMEP should be prioritized in single-chemical models
2. **Mixture models:** Include URXCOP, URXMPB, URXEPB alongside URXMEP
3. **Essential covariates:** Smoking status, reproductive history, age
4. **Model comparison:** Compare linear vs. non-linear approaches given the discrepancy between methods
5. **Stratification:** Consider stratifying by smoking status or reproductive history

### Next Steps
- Build regression models with URXMEP as primary exposure
- Test mixture effects using WQS or BKMR
- Stratify analyses by smoking status
- Investigate why Random Forest didn't select URXMEP
- Compare model performance (linear vs. non-linear)


# Summary of Feature Selection Results for SHBG (Sex Hormone-Binding Globulin)

## Method Performance

### LASSO Regression
- **Selected features:** 16 features
- **Performance:** Much more successful than for estradiol (which selected 0 features)
- **Interpretation:** Indicates stronger linear relationships between predictors and SHBG

### Elastic Net (alpha = 0.5)
- **Selected features:** 16 features (same as LASSO)
- **Interpretation:** High agreement with LASSO suggests stable feature selection

### Random Forest
- **Selected features:** 32 features with positive importance
- **Performance:** Best performing method in terms of identifying important variables
- **Note:** Shows model performance metrics with R-squared

### Correlation Analysis
- **Selected features:** 31 features with |r| > 0.05
- **Interpretation:** Identified features with strongest linear relationships

---

## Top Correlations with SHBG

| Variable | Description | Correlation (r) | Direction |
|----------|-------------|-----------------|-----------|
| **BMXBMI** | **Body Mass Index** | **-0.271** | **Negative** |
| RHQ131 | Ever been pregnant | 0.179 | Positive |
| SMQ040 | Smoking status | 0.153 | Positive |
| DMDEDUC2 | Education level | 0.146 | Positive |
| RIDAGEYR | Age | -0.144 | Negative |
| log_URXMBP | Mono-n-butyl phthalate | -0.143 | Negative |
| RHQ074 | Pregnancy status | 0.140 | Positive |
| log_URX14D | 2,5-Dichlorophenol | -0.135 | Negative |
| log_URXMIB | Mono-isobutyl phthalate | -0.134 | Negative |
| INDFMPIR | Income-to-poverty ratio | 0.132 | Positive |

---

## Feature Selection by Method Consensus

### Consensus Features (Selected by ALL 4 Methods) - **12 Features**

**Reproductive Health:**
- `RHQ074` - Pregnancy status

**Lifestyle Factors:**
- `SMQ040` - Smoking status
- `ALQ120Q` - Alcohol consumption

**Socioeconomic Status:**
- `INDFMPIR` - Poverty income ratio
- `DMDEDUC2` - Education level

**Anthropometric:**
- `BMXBMI` - Body Mass Index

**Demographic:**
- `RIDAGEYR` - Age

**EDC Chemicals:**
- `log_URXBUP` - Butyl paraben
- `log_URXMIB` - Mono-isobutyl phthalate
- `log_URXBPH` - Bisphenol A
- `log_URXCOP` - Mono(carboxyoctyl) phthalate
- `log_URX14D` - 2,5-Dichlorophenol

### Features Selected by ≥3 Methods - **16 Features**

All 12 consensus features above, **plus:**
- `RHQ031` - Regular menstrual periods
- `log_URXMHNC` - Cyclohexane-1,2-dicarboxylic acid
- `RHQ131` - Ever been pregnant
- `RHD180` - Age at first live birth

### Features Selected by ≥2 Methods - **29 Features**

All 16 features above, **plus 13 additional** phthalate metabolites and BPA compounds (Random Forest + Correlation):
- `log_URXDCB`, `URXUCR`, `log_URXMHH`, `log_SSURMHBP`, `log_URXMBP`
- `log_URXMOH`, `log_URXECP`, `log_URXMZP`, `log_URXMPB`, `log_URXMHP`
- `log_SSURHIBP`, `log_URXBPS`, `log_URXBPF`

---

## Key EDC Chemicals Associated with SHBG

### Strong Negative Associations
*Lower chemical exposure → Higher SHBG*

**Phthalate Metabolites:**
- `log_URXMBP` (Mono-n-butyl phthalate): r = -0.143
- `log_URXMIB` (Mono-isobutyl phthalate): r = -0.134
- `log_URXECP` (Mono(2-ethyl-5-carboxypentyl) phthalate): r = -0.117

**Phenols:**
- `log_URX14D` (2,5-Dichlorophenol): r = -0.135

**Bisphenols:**
- `log_URXBPH` (Bisphenol A): r = -0.117
- `log_SSURHIBP` (BPA metabolite): r = -0.108
- `log_URXBPS`, `log_URXBPF` (BPA alternatives)

**Parabens:**
- `log_URXBUP` (Butyl paraben)

---

## Important Non-EDC Covariates

### Strongest Predictor: BMI
- **Correlation:** r = -0.271 (by far the strongest)
- **Relationship:** Higher BMI → Lower SHBG
- **Clinical significance:** Well-established relationship in literature
- **Implication:** Must be carefully controlled in all models

### Lifestyle Factors
- **Smoking (SMQ040):** r = 0.153
  - Relationship: Smokers have higher SHBG (positive association)
- **Alcohol (ALQ120Q):** Selected by all 4 methods

### Socioeconomic Status
- **Education (DMDEDUC2):** r = 0.146
  - Relationship: Higher education → Higher SHBG
- **Income (INDFMPIR):** r = 0.132
  - Relationship: Higher income → Higher SHBG

### Reproductive History
All strongly associated with SHBG:
- `RHQ074` (Pregnancy status)
- `RHQ131` (Ever been pregnant)
- `RHQ031` (Regular menstrual periods)
- `RHD180` (Age at first live birth)

### Age
- **RIDAGEYR:** r = -0.144
  - Relationship: Younger age → Higher SHBG

---

## Overall Interpretation

> **SHBG shows the STRONGEST and most consistent relationships with both EDCs and covariates compared to testosterone and estradiol.**

### Key Observations

#### 1. Excellent Method Agreement
- **12 features selected by ALL 4 methods** (vs. 0 for estradiol)
- Both linear and non-linear methods agree
- Suggests robust, interpretable relationships

#### 2. BMI Dominates
- Nearly **double the correlation** of any other variable (r = -0.271)
- Must be carefully controlled in all models
- May mediate EDC effects on SHBG

#### 3. Multiple EDCs Matter
- Unlike testosterone (dominated by URXMEP alone), SHBG shows associations with **many different EDCs**
- Suggests **mixture effects are important**
- Phthalates, parabens, and BPA all relevant

#### 4. Socioeconomic Factors
- Education and income positively associated with SHBG
- May reflect lifestyle, diet, or exposure patterns
- Important confounders to adjust for

#### 5. Negative EDC Associations
- Most EDCs **negatively correlated** with SHBG
- **Clinical significance:** Lower SHBG → More bioavailable hormones → Potential health effects
- Consistent direction across multiple chemical classes

---

## Comparison Across Hormones

| Hormone | Consensus Features | Strongest EDC | Top Correlation | Model Performance |
|---------|-------------------|---------------|-----------------|-------------------|
| **SHBG** | **12** | Multiple | r = -0.271 (BMI) | **Best** |
| Testosterone | 0-3 | URXMEP | r = 0.19 (URXMEP) | Moderate |
| Estradiol | 0 | None clear | r = 0.158 (Age) | Poor |

**SHBG emerges as the best outcome for EDC research.**

---

## Implications for Modeling

### Critical Considerations

1. **BMI is critical**
   - Must include in all models
   - Consider stratification by BMI categories
   - May modify EDC effects

2. **Mixture models essential**
   - Use WQS or BKMR to capture combined EDC effects
   - Multiple chemicals matter, not just single exposures

3. **Linear models appropriate**
   - Strong agreement between LASSO/Elastic Net and Random Forest
   - Linear relationships work well for SHBG

4. **Include SES covariates**
   - Education and income are important confounders
   - May capture unmeasured lifestyle/dietary factors

5. **Best outcome for EDC research**
   - SHBG appears most responsive to EDC exposures
   - Clearest, most consistent signals

---

## Recommended Model Structure

### Primary Outcome
- **SHBG** (most promising for detecting EDC effects)

### Exposure Variables
- **EDC mixture** including:
  - Multiple phthalate metabolites (URXMBP, URXMIB, URXECP, etc.)
  - Bisphenol A and alternatives (URXBPH, URXBPS, URXBPF)
  - Parabens (URXBUP)
  - Phenols (URX14D)

### Essential Covariates
- **BMI** (strongest predictor)
- **Age** (RIDAGEYR)
- **Smoking** (SMQ040)
- **Education** (DMDEDUC2)
- **Income** (INDFMPIR)
- **Reproductive history** (RHQ074, RHQ131, RHQ031, RHD180)

### Stratification Considerations
- **By BMI categories** (normal, overweight, obese)
  - Test whether BMI modifies EDC-SHBG associations
- **By reproductive status**
  - Nulliparous vs. parous women

---

## Next Steps

1. **Prioritize SHBG** as primary outcome for EDC mixture analysis

2. **Build comprehensive models**
   - Include all 12 consensus features as baseline
   - Test additional features from 3-method and 2-method lists

3. **Implement mixture analysis**
   - WQS regression to estimate mixture effects
   - BKMR to assess non-linear dose-response and interactions

4. **Stratification analyses**
   - By BMI categories to test effect modification
   - By smoking status (given strong association)

5. **Comparative analyses**
   - Compare SHBG models with testosterone and estradiol
   - Understand hormone-specific EDC effects
   - Determine which hormones are most sensitive to environmental exposures

6. **Sensitivity analyses**
   - Adjust for creatinine vs. creatinine-adjusted EDC concentrations
   - Exclude extreme BMI values
   - Restrict to specific subpopulations

7. **Mechanistic exploration**
   - Investigate whether BMI mediates or moderates EDC effects
   - Explore pathways: EDCs → SHBG → bioavailable hormones



