---
title: "playground"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(knitr)
library(haven)
library(dplyr)
library(car)
library(survey)
library(srvyr)
```

## Read datasets

```{r}
#Nacist data a z nekterych vybrat jen nektere promenne

#Demographics, RIDAGEYR - vek, RIDEXPRG - pregnant, DMDEDUC2 - education level, RIAGENDR - gender, INDFMPIR - ratio of family income to poverty 
#RIDSTATR - both interview and MEC examination (2 - both, 1 - only interview)
demo_data = read_xpt("./data/DEMO_H.xpt") |> select(SEQN,RIDAGEYR,RIDEXPRG,DMDEDUC2,RIAGENDR,INDFMPIR,WTMEC2YR, SDMVPSU,SDMVSTRA)
#Reproductive health data, RHQ031 - had regular periods in past 12 months, RHQ074 - tried for a yer to become pregnant, RHQ076 - seen DR unable to become pregnant, RHQ131 - ever been pregnant, 
#RHQ160 - how many times have been pregnant, RHQ171 - live birth deliveries, RHD280 - had a hysterectomy, RHQ078 - pelvic infection,RHD180 - age at firts livebirth
rhq_data = read_xpt("./data/RHQ_H.xpt") |> select(SEQN,RHQ031,RHQ074,RHQ076, RHQ131,RHQ160, RHQ171, RHD280, RHQ078, RHD180)
# SEx steroide hormone, LBXTST - testosterone (ng/dL), LBXEXT - estradiol(pg/mL), LBXSHGB (nmol/L), promene koncici na LC (0 - above detection limit, 1 below detection limit)
tst_data = read_xpt("./data/TST_H.xpt")
#Phtalates and plasticizesr metabolites
ssphte_data = read_xpt("./data/SSPHTE_H.xpt")

#Personal care and consumer product chemicals and metabolites
ephpp_data = read_xpt("./data/EPHPP_H.xpt")

#phtalates and plasticizers metabolites
phthte_data = read_xpt("./data/PHTHTE_H.xpt")

#alcohol use, ALQ120Q - How often drink alcohol over past 12 months
alq_data = read_xpt("./data/ALQ_H.xpt") |> select(SEQN,ALQ120Q)
#smoking cigarette use, SMQ040 - do you now smoke cigarettes, SMD650 avg cigarettes a day last month
smq_data = read_xpt("./data/SMQ_H.xpt") |> select(SEQN,SMQ040,SMD650)
#medical condition
#mcq_data = read_xpt("./data/MCQ_H.xpt") |> select(SEQN,)
#body measures, BMXBMI - body maxx index
bmx_data = read_xpt("./data/BMX_H.xpt") |> select(SEQN,BMXBMI)


```

## Merging datasets

```{r}
#iiner join keep only rows where there is a SEQN match in both datasets
merged_data = inner_join(demo_data, rhq_data, by = "SEQN")
merged_data = inner_join(merged_data, tst_data, by = "SEQN")
merged_data = inner_join(merged_data, ssphte_data, by = "SEQN")
merged_data = inner_join(merged_data, ephpp_data, by = "SEQN")
merged_data = inner_join(merged_data, alq_data, by = "SEQN")
merged_data = inner_join(merged_data, smq_data, by = "SEQN")
merged_data = inner_join(merged_data, bmx_data, by = "SEQN")
merged_data = inner_join(merged_data, phthte_data, by = "SEQN")
nrow(merged_data)
#jedna promena se vyskytuje ve vice datasetech
merged_data <- merged_data[, !grepl("\\.y$", names(merged_data))]
names(merged_data)[names(merged_data) == "URXUCR.x"] <- "URXUCR"
names(merged_data)[names(merged_data) == "WTSB2YR.x"] <- "WTSB2YR"

var_labels <- sapply(merged_data, function(x) attr(x, "label"))
labels_df <- data.frame(
  variable = names(var_labels),
  label = unname(var_labels)
)
print(labels_df)
```

## Remove, based on exclusion, inclusion criteria from the plan

```{r}
#age between 18 and 45 and female only, not pregnant
clean_data <- merged_data[merged_data$RIDAGEYR >= 18 & merged_data$RIDAGEYR <= 50 & merged_data$RIAGENDR == 2 & merged_data$RIDEXPRG == 2,]
nrow(clean_data)
#extreme urinary creatine values

clean_data <- clean_data[clean_data$URXUCR<300,]

#missing key covariates -> R si NaN values odstraní automaticky pri vytvareni modelu ?

#urinary chemicals measurement and MEC examination
clean_data <- clean_data[clean_data$WTSSBH2Y>0 & clean_data$WTSB2YR>0 & clean_data$WTMEC2YR>0,]
nrow(clean_data)

#at least one detectable urinary chemical measurement (ze vsech chemikalii v EPHPP a SSPHTE)
all_vars <- union(union(names(ssphte_data), names(ephpp_data)),names(phthte_data))
exclude_vars <- c( "SEQN","WTSSBH2Y","WTSB2YR" )
edc_vars <- setdiff(all_vars, exclude_vars)
length(edc_vars)
data_1 <- clean_data[
  rowSums(!is.na(clean_data[, edc_vars])) >= 1, #alespon jedna hodnota ze vsech merenych chemikalii neni Nan (missing)
]
# at least one detectable hormone measurement (not missing - nan)
horm_vars <- c("LBXSHBG","LBXEST","LBXTST")
data_1 <- data_1[
  rowSums(!is.na(data_1[, horm_vars])) >= 1, #alespon jedna hodnota ze vsech hormonu neni nan
]

nrow(data_1)
```

## Survey design object -> to work with the weights
NHANES: You must use the weight of the smallest subpopulation that includes all the variables you want to include in your analysis.
Z tech mensich souboru obsahují speciální vahy jenom ty chemikálie (EPHPP a SSPHTE a PHTHTE). EPHPP a PHTHTE maji stejne (WTSB2YR) a SSPHTE ma jiné (WTSSBH2Y). Nevim ktera z nich by se mela pouzit, ale nejmensi je dataset SSPHTE.

```{r}
# v ukazkach pak na vsechny upravy pouzivaji uz jenom survey desing objekt -> vytvorit z celych dat a znovu stejne profiltrovat
data_w <- svydesign(id = ~SDMVPSU, strata= ~SDMVSTRA, weight=~WTSB2YR,   nest=TRUE, data=merged_data)

obj_cleaned <- data_w |>
  subset(
    RIDAGEYR >= 18 &
    RIDAGEYR <= 50 &
    RIAGENDR == 2 &
    RIDEXPRG == 2 &
    URXUCR < 300 &
    WTSSBH2Y > 0 &
    WTSB2YR > 0 &
    WTMEC2YR > 0 &
    rowSums(!is.na(as.data.frame(data_w$variables[, edc_vars]))) >= 1 &
    rowSums(!is.na(as.data.frame(data_w$variables[, horm_vars]))) >= 1
  )
nrow(obj_cleaned$variables)
summary(obj_cleaned)
#vytvoreny survey objekt porad obsahuje sloupce s vahama z vetsich datasetu -> pri vytvareni modelu upravit a nebrat je v uvahu
```

##  Variables statistics

```{r}
#age
svymean(~RIDAGEYR,obj_cleaned,na.rm = TRUE)


```
```{r}
 #visualize
svyhist(~URXBPH, design=obj_cleaned, main="Bisphenol A concentration")

svyhist(~log(URXBPH + 1), design=obj_cleaned, main="Bisphenol A concentration, log transformed")
```
```{r}
#Correlation matrix - na vsem krome chemikalii, weights a gender a pregnant, protoze to jsou po filtraci vlastne konstantni sloupce pro vsechny stejne
exclude_vars <- c("SEQN","WTMEC2YR", "WTSSBH2Y", "WTSB2YR",edc_vars,"SDMVPSU","SDMVSTRA","RIDEXPRG","RIAGENDR")
cor_vars <- setdiff(names(obj_cleaned$variables), exclude_vars)
length(cor_vars)

cor_matrix <- cor(obj_cleaned$variables[, cor_vars],use = "pairwise.complete.obs")
corrplot::corrplot(cor_matrix,tl.cex = 0.6,method = "color",type = "upper")


```
```{r}
# multicollinearity check
# correlation between the chemicals (exposures), phthalates and their metabolites could be significantly correlated
phthalate_vars <- c('URXUCR', 'SSURMHBP', 'SSURHIBP', 'URXCNP', 'URXCOP', 'URXECP', 'URXMBP', 'URXMC1', 'URXMEP', 'URXMHH', 'URXMHNC', 'URXMHP', 'URXMIB', 'URXMNP', 'URXMOH', 'URXMZP')
phenol_vars <- c('URXBP3', 'URXBPH', 'URXBPF', 'URXBPS', 'URXTLC', 'URXTRS', 'URXBUP', 'URXEPB', 'URXMPB', 'URXPPB', 'URX14D', 'URXDCB')

chem_vars <- c(phthalate_vars, phenol_vars)

chem_data <- obj_cleaned$variables[, chem_vars]
cor_mat <- cor(chem_data, use = "pairwise.complete.obs")
corrplot::corrplot(cor_mat,tl.cex = 0.6,method = "color",type = "upper")

# VIF (Variance inflation factor)
# VIF > 5: high, VIF > 10: severe multicollinearity


```
```{r}
# EDC concentrations are right-skewed, before PCA log-transform the chemicals
chem_log <- log(chem_data)
# standardizing, scaling
chem_scaled <- scale(chem_log)

pca_res <- prcomp(chem_scaled, center = TRUE, scale. = TRUE)
summary(pca_res) # explained variance - how many PCs to keep?

# how many PCs to keep? (practical rules)
# 1.) scree plot ("elbow rule") most common
# look for clear elbow, keep PCs before the curve flattens
plot(pca_res, type = "l")

# 2.) variance threshold - common threshold ~ 50-70% cummulative variance
model <- svyglm(
  log(LBXTST) ~ URXECP + URXMHH + URXCOP + URXMNP + RIDAGEYR + BMXBMI, # gender doesn't make sense, we have just women
  design = obj_cleaned
)
vif_values <- vif(model)

# View the results
print(vif_values)

```

```{r}
loadings <- pca_res$rotation
print(loadings)
```



```{r}

# typically: PC1 = overall exposure burden, PC2 = different metabolite patterns
# add PCA scores back to your survey object
pc_scores <- as.data.frame(pca_res$x[, 1:2])  # first 2 PCs
colnames(pc_scores) <- c("PC1", "PC2")
obj_cleaned$variables <- cbind(obj_cleaned$variables, pc_scores)
# PC1 and PC2 can be used as exposure variables 
```

```{r}
obj_cleaned$variables$SMQ040

```


```{r}
# survey weighted regression using PCA scores
# testosterone
# factor() - category, not a number
obj_cleaned <- update(obj_cleaned, 
                      SMOKE_STATUS = factor(SMQ040, 
                                            levels = c(3, 2, 1), 
                                            labels = c("Never", "Some_Days", "Every_Day")))

model <- svyglm(
  log(LBXTST) ~ PC1 + RIDAGEYR + SMOKE_STATUS, # handle SMQ040
  design = obj_cleaned
)

summary(model)
# checking the assumptions
# multicollinearity - ok (PCA)
# normality of residuals - ok
qqnorm(residuals(model))
qqline(residuals(model))
# heteroscedasticity - formal tests (like Breusch-Pagan) are not appropriate for survey models
# svyglm() uses robust standard errors - mild heteroscedasticity is acceptable
# roughly symmetric around zero, no clear patterns
# also linearity check?
plot(fitted(model), residuals(model))
# data distribution, outliers
boxplot(log(obj_cleaned$variables$LBXTST))
x <- log(obj_cleaned$variables$LBXTST) # outliers flagged by the boxplot rule (> Q3 + 1.5×IQR or < Q1 − 1.5×IQR)

bp <- boxplot.stats(x)
bp$out
exp(bp$out)

confint(model)

r_dev <- residuals(model, type = "deviance")
plot(fitted(model), r_dev) # linear structure, we want no pattern or curvature
abline(h = 0, lty = 2)

# normality of the error distribution is not required for consistency
qqnorm(r_dev)
qqline(r_dev) # mild tail deviations are ok

nrow(model$model)

```
```{r}
library(splines)
# estradiol
model <- svyglm(
  log(LBXEST) ~ PC1 + RIDAGEYR + BMXBMI,
  design = obj_cleaned
)

summary(model)
confint(model)

r_dev <- residuals(model, type = "deviance")
plot(fitted(model), r_dev) # linear structure, we want no pattern or curvature
abline(h = 0, lty = 2)

# normality of the error distribution is not required for consistency
qqnorm(r_dev)
qqline(r_dev) # mild tail deviations are ok

# heteroscedasticity - the weights should deal with that
plot(fitted(model), abs(r_dev))

boxplot(log(obj_cleaned$variables$LBXEST))
x <- log(obj_cleaned$variables$LBXEST)

bp <- boxplot.stats(x)
bp$out
exp(bp$out)

influence <- influence(model)
summary(influence$hat) # leverage
summary(influence$cooks) # influence

# weight distribution
hist(weights(model), breaks = 50)
summary(weights(model))

# multicollinearity
vif(model)

nrow(model$model)


```
```{r}
model1 <- svyglm(log(LBXEST) ~ PC1 + RIDAGEYR + BMXBMI, design = obj_cleaned)
model2 <- svyglm(log(LBXEST) ~ PC1 + ns(RIDAGEYR,3) + BMXBMI, design = obj_cleaned)
model3 <- svyglm(log(LBXEST) ~ PC1 + RIDAGEYR + ns(BMXBMI,3), design = obj_cleaned)

summary(model1)
confint(model1)
summary(model2)
confint(model2)
summary(model3)
confint(model3)
```

```{r}
# level of education DMDEDUC2
# 1 ... less than 9th grade
# 2 ... 9-11th grade
# 3 ... high school graduate/GED or equivalent
# 4 ... some college or AA degree
# 5 ... college graduate or above
# 7 ... refused

obj_cleaned$variables$DMDEDUC2
```

```{r}
# alcohol consumption
# ALQ120Q: How often drink alcohol over past 12 months?, range 0 to 365

obj_cleaned$variables$ALQ120Q
# no 777 or 999 ... no need to handle it

```

```{r}
# ratio of family income to poverty INDFMPIR
# continuous, 0 to 5

obj_cleaned$variables$INDFMPIR
```

```{r}
# ratio of family income to poverty INDFMPIR
# continuous, 0 to 5

obj_cleaned$variables$SM
```



```{r}

# SHBG
# age, bmi, smoking, alcohol, education, income
# strongest predictor should be BMI, well-known strong inverse relationship between BMI and SHBG
# alcohol consumption often increases SHBG levels
obj_cleaned <- update(obj_cleaned, 
    # 1. Education as a Factor
    # We filter out 7 (Refused) and 9 (Don't Know) to keep the analysis clean
    EDUCATION_LEVEL = factor(ifelse(DMDEDUC2 <= 5, DMDEDUC2, NA),
                        levels = c(1, 2, 3, 4, 5),
                        labels = c("Less_than_9th", "9-11th", "HS_Grad", "Some_College", "College_Grad")))
    
# how to decide on the confounders?
# -> look at the coeffiecient of my main predictors (here BMI and PC1)
# if dropping a variable changes those coefficients by less than 10%, the variable is likely not a confounder
# gain in statistical power - we are increasing our sample size


model <- svyglm(
  #log(LBXSHBG) ~ PC1 + RIDAGEYR + BMXBMI + SMOKE_STATUS + ALQ120Q + EDUCATION_LEVEL + INDFMPIR,
  #log(LBXSHBG) ~ PC1 + RIDAGEYR + BMXBMI + SMOKE_STATUS + EDUCATION_LEVEL + INDFMPIR,
  #log(LBXSHBG) ~ PC1 + RIDAGEYR + BMXBMI + EDUCATION_LEVEL + ALQ120Q + INDFMPIR,
  log(LBXSHBG) ~ PC1 + RIDAGEYR + BMXBMI + EDUCATION_LEVEL + INDFMPIR,
  design = obj_cleaned
)

summary(model)
confint(model)  # doesn't work here, probably because the model is too sparse, big part of the data got dropped due to NaN values

r_dev <- residuals(model, type = "deviance")
plot(fitted(model), r_dev) # linear structure, we want no pattern or curvature
abline(h = 0, lty = 2)

# normality of the error distribution is not required for consistency
qqnorm(r_dev)
qqline(r_dev) # mild tail deviations are ok

boxplot(log(obj_cleaned$variables$LBXSHBG))
x <- log(obj_cleaned$variables$LBXSHBG)

bp <- boxplot.stats(x)
bp$out
exp(bp$out)

# multicollinearity
check_vif <- lm(log(LBXSHBG) ~ PC1 + RIDAGEYR + BMXBMI + SMOKE_STATUS + ALQ120Q + EDUCATION_LEVEL + INDFMPIR, 
                data = obj_cleaned$variables)
vif(check_vif)

nrow(model$model) # only 71 observations x 324 (a lot gets dropped due to NaN values)

```

```{r}
# what has the most NA values
variables_to_check <- c("LBXSHBG", "PC1", "RIDAGEYR", "BMXBMI", 
                        "SMQ040", "ALQ120Q", "DMDEDUC2", "INDFMPIR")

sapply(obj_cleaned$variables[variables_to_check], function(x) sum(is.na(x)))
```


```{r}
# sanity check
apply(chem_scaled, 2, mean) # should be ~0
apply(chem_scaled, 2, sd) # should be ~1
```




```{r}



```


